A Pilot Multicenter Cluster Randomized Trial to Compare The Effect of Trauma Life Support Training Programs on Patient and Provider Outcomes

### 2022-12-06 pilot-publication/manuscript.Rmd
Update analysis with 235 patients

## Study Setting
We conducted this pilot in seven Indian tertiary hospitals, where
neither ATLS, PTC, nor any other trauma life support training program
is routinely taught. The seven hospitals were King Edward Memorial
Seth G. S. Medical College and K.E.M. Hospital, Mumbai; Lokmanya Tilak
Municipal Medical College and General Hospital, Mumbai; HBT Medical
College And Dr. R N Cooper Municipal General Hospital, Mumbai; Medica
Superspecialty Hospital, Kolkata; Medical College Kolkata, Kolkata;
Sir Nil Ratan Sircar Medical College & Hospital, Kolkata; Postgraduate
Institute of Medical Education & Research, Chandigarh.

@@

### Patient Participants
Adults (15 years or older) who present to the emergency department at
participating hospitals with a history of trauma. History of trauma is
here defined as having any of the external causes of morbidity and
mortality listed in block V01-Y36, chapter XX of the International
Classification of Disease version 10 (ICD-10) codebook as reason for
presenting. We explored intervention effects across the following
clinical subgroups: men, women, blunt multisystem trauma, penetrating
trauma, shock, severe traumatic brain injury, and elderly, as defined
by Hornor et al [@Hornor2018]. The consent form for patients are
available as Supplemental Material 1.

@@

### Clusters
Indian tertiary care hospitals that admit 400-800 adult patients with
trauma each year. We randomised on the cluster (hospital) level to
avoid contamination between intervention and control arms. To be
eligible for inclusion hospitals had to provide the following services
round the clock: operation theatres, X-ray, CT, and ultrasound
facilities, and blood bank. In addition the baseline admission rate
had to be more than 35 adult patients with major trauma per month.

@@

## Patient Participant Outcomes
In the post training period, a total of `r n.m30d.atls` (
`r p.m30d.atls`%) patients in the ATLS arm had the primary outcome,
compared to `r n.m30d.ptc` (`r p.m30d.ptc`%) patients in the PTC arm,
and `r n.m30d.control` (`r p.m30d.control`%) patients in the standard
care arm. The absolute risk difference between the ATLS and PTC arm
was `r arr.atls.ptc`% units <!--(95% CI X - X)-->, between the ATLS and 
standard care arms `r arr.atls.control`% units <!--(95% CI X - X)-->, and 
between the PTC and standard care arms `r arr.ptc.control`% units 
<!--(95% CI X -X)-->. The relative risk between the ATLS and 
PTC arm was `r rr.atls.ptc` <!--(95% CI X - X)-->, between the ATLS and
standard care arms `r rr.atls.control` <!--(95% CI X - X)-->, and 
between the PTC and standard care arms `r rr.ptc.control` <!--(95% CI X - X)-->.

@@

Differences in secondary outcomes between trial arms and subgroups are
available as Supplementary material \@ref(tab:secondary-outcomes).

@@

## Feasibility Outcomes
<!--Out of X potentially eligible patients, X (X%) patients were enrolled
in the study-->. Out of `r n.eligible.residents` potentially eligible
residents, `r n.residents` (`r recruitment.rate.residents`%) were
enrolled in the study. The lost to follow up rate was 
`r rate.lost.to.follow.up`, with `r n.lost.to.follow.up` out of 
`r n.patients` patients not completing 30 day follow up. Among residents
the pass rate, after two attempts, was `r rate.pass`%. The missing
data rate ranged from `r min.missing.data` to `r max.missing.data`,
with details per variable and outcome in Table X. <!--The differences in
distributions between observed and extracted data, for variables that
were collected through observation or interview, are shown in Table X.-->

>>>>

## Study Setting
We conducted this pilot in seven Indian tertiary hospitals, where
neither ATLS, PTC, nor any other trauma life support training program
is routinely taught. The seven hospitals were Seth G. S. Medical
College and K.E.M. Hospital, Mumbai; Lokmanya Tilak Municipal Medical
College and General Hospital, Mumbai; HBT Medical College And Dr. R N
Cooper Municipal General Hospital, Mumbai; Medica Superspecialty
Hospital, Kolkata; Medical College Kolkata, Kolkata; Sir Nil Ratan
Sircar Medical College & Hospital, Kolkata; Postgraduate Institute of
Medical Education & Research, Chandigarh.

@@

### Patient Participants
Adults (15 years or older) who presented to the emergency department at
participating hospitals with a history of trauma. History of trauma is
here defined as having any of the external causes of morbidity and
mortality listed in block V01-Y36, chapter XX of the International
Classification of Disease version 10 (ICD-10) codebook as reason for
presenting. We explored intervention effects across the following
clinical subgroups: men, women, blunt multisystem trauma, penetrating
trauma, shock, severe traumatic brain injury, and elderly, as defined
by Hornor et al [@Hornor2018]. The consent form for patients are
available as Supplemental Material 1.

@@

### Clusters
Indian tertiary care hospitals that admit more than 400 adult patients
with trauma each year. We randomised on the cluster (hospital) level
to avoid contamination between intervention and control arms. To be
eligible for inclusion hospitals had to provide the following services
round the clock: operation theatres, X-ray, CT, and ultrasound
facilities, and blood bank. In addition the baseline admission rate
had to be more than 35 adult patients with major trauma per month.

@@

## Patient Participant Outcomes
A total of `r n.m30d.atls` ( `r p.m30d.atls`%) patients in the ATLS
arm had the primary outcome, compared to `r n.m30d.ptc` (
`r p.m30d.ptc`%) patients in the PTC arm, and `r n.m30d.control` (
`r p.m30d.control`%) patients in the standard care arm. The absolute risk
difference between the ATLS and PTC arm was `r arr.atls.ptc`% units
<!--(95% CI X - X)-->, between the ATLS and standard care arms 
`r arr.atls.control`% units <!--(95% CI X - X)-->, and between the PTC
and standard care arms `r arr.ptc.control`% units <!--(95% CI X
-X)-->. The relative risk between the ATLS and PTC arm was 
`r rr.atls.ptc` <!--(95% CI X - X)-->, between the ATLS and standard care
arms `r rr.atls.control` <!--(95% CI X - X)-->, and between the PTC
and standard care arms `r rr.ptc.control` <!--(95% CI X - X)-->.

@@

Differences in secondary outcomes between trial arms<!-- and subgroups--> are
available as Supplementary material table \@ref(tab:secondary-outcomes).

@@

## Feasibility Outcomes
<!--Out of X potentially eligible patients, X (X%) patients were enrolled
in the study-->. Out of `r n.eligible.residents` potentially eligible
residents, `r n.residents` (`r recruitment.rate.residents`%) were
enrolled in the study. The lost to follow up rate was 
`r rate.lost.to.follow.up`, with `r n.lost.to.follow.up` out of 
`r n.patients` patients not completing 30 day follow up. Among residents
the pass rate, after two attempts, was `r rate.pass`%. The missing
data rate ranged from `r min.missing.data` to `r max.missing.data`,
with details per variable and outcome in Table 
\ref@(tab:sample-characteristics) and Supplementary material table
\@ref(tab:secondary-outcomes). <!--The differences in
distributions between observed and extracted data, for variables that
were collected through observation or interview, are shown in Table X.-->
